\chapter{Related Work and Background}\label{ch_background}

Life Fitness in 2012 released a survey concerning technology in the
exercise environment\cite{lifefitness} where it was found that over
half of gym users used a smart phone or tablet when exercising. This
study focused solely on gym users which is not our target market for
this product (we are targeting users who exercise outdoors), however
it is not unreasonable to assume similar statistics for our target
demographic. We use this metric along with the vast user base of the
related applications we compare in Section \ref{sec_comparison} as a
basis of platform validity. 

\section{Gamification}
\citet{gamification_book} define gamification as ``The process of
game-thinking and game mechanics to engage users and solve
problems.''. There are four main building blocks to this definition:

\begin{description}[style=multiline, leftmargin=3cm]
\item[\emph{game-thinking}] What the users thoughts should be when
  they are interacting with this game and how the user should perceive
  their interaction.
\item[\emph{game mechanics}] How we provide this game-like
  experience. What metrics we use to indicate success and encourage
  play. 
\item[\emph{users}] What user group this gamification is intended for
  and who we hope to influence the most through this application.
\item[\emph{problems}] What issues are we trying to fix or improve
  through this application of gamification.
\end{description}

\citet{artOfGameDesign} proposes four main elements, the ``elemental
tetrad'', that form a game: \emph{mechanics}, \emph{story},
\emph{aesthetics}, and \emph{technology}. These four elements influence
how a user perceives a game (\emph{game-thinking}) and
\citeauthor{artOfGameDesign} proposes that all four should be treated with
equal weighting to achieve a successful game. An engrossing
\emph{story} combined with good \emph{aesthetics} can be used to mask
what a user is truly doing. Recently, \citet{genesInSpace}
released a game called \emph{Genes In Space} where users chart a
course through space to pick up as much ``Element Alpha'' as they can
while destroying asteroids. The route they chart is based on a space
map where the distribution of ``Element Alpha'' is shown and
collecting ``Element Alpha'' allows the user to buy upgrades to their
spaceship to personalise and improve it. The distribution of ``Element
Alpha'' is provided by a DNA microarray of a cancer patient and the 
course you plot identifies where alterations in the genome have
occurred. By applying game elements the user plots a course that will
maximise their success (\emph{game mechanic}) while they are also
identifying key areas of interest that will benefit cancer
researchers. 

A similar application for protein folding is utilised by the Foldit
project by \citet{foldit}. Here users are challenged directly to
manipulate a protein chain to find an optimal layout of this chain (a
fold) based on a set of given criteria. This solution allows
suggestions to be proposed by the community and reviewed by scientists
to determine their merit. There is a points based \emph{game mechanic}
where optimal layouts are rewarded with more points than less optimal
solutions. A benefit to this application of gamification is that it
saves a substantial amount of the scientists time. The scientist only
has to review the submitted protein fold which is a much shorter time
investment than designing the fold.

In both of the applications of gamification above the \emph{problems}
have been of a scientific nature and the \emph{users} have been casual
gamers, scientists and researchers. Gamification has been shown to
bring improvements in domains other than medical research.

\emph{World Without
  Oil}\footnote{\url{http://www.worldwithoutoil.org}} was an Alternate
Reality Game created to explore the ways to conserve oil resources
during a shortage where 1,700 players were involved in the project
over a 32 day period. \citet{reality_broken}, one of the designers
involved with \emph{World Without Oil}, commented about the
behavioural change benefits of games:

\begin{quote}
  We can change our real-life
  behaviours in the context of a fictional game precisely because there
  isn't any negative pressure surrounding the decision to change. We
  are motivated purely by positive stress and by our own desire to
  engage with a game in more satisfying, successful, social and
  meaningful ways.
\end{quote}

The \emph{problems} here were based on a contrived situation, but
\citeauthor{reality_broken} notes that when there was a gas crisis a
year later players of \emph{World Without Oil} were in a significantly
better situation than those who were not. This application of
gamification allowed these users to positively benefit in a real world
situation that was similar to one experienced in game.

A final example closer to the problem domain of this project is
\emph{Chick Clique} (\citet{chick}) which uses social game elements to
encourage a peer group of four teenage schoolgirls to exercise. Each
student was given a personal digital assistant (PDA) and a pedometer
to track the steps they take. The steps taken can be compared with the
other three peers in their group to gauge personal progress. Help was
also provided where the user could get information about fast-food in
terms of how many steps were required to ``walk off'' the empty
calories. Although this was a small experiment it was met with good
feedback from the test group who commented that breaching the subject
of healthy eating and exercise within the group was as valuable as the
tracking and suggestion the product gave. This is another example
where gamification has positive side effects outside the game
environment.

\section{Comparison of related applications}\label{sec_comparison}
Incentivising exercise is not a new concept in the mobile market. From
simple applications that mimic a pedometer to immersive alternate
realities, there has been a varied approach to encouraging
exercise. These approaches can be shown through the following
applications. 

Table \ref{table:competitor_comparison} gives a brief
overview of requirements and user statistics for the compared products
gained from the Google Play Store.

\begin{table}[h]
  \centering
  \begin{tabular}{ | c | c | c | c |} \hline
    Name & Interface & Requirements & User base  \\ \hline
    Charity Miles & Mobile App & App and GPS & 10K - 50K\cite{charitymiles_store}\\ \hline
    Zombies, Run! & Mobile App & App and GPS, headphones & 100K - 500K\cite{zombiesrun_store}\\ \hline
    WalkJogRun & Mobile App & App and GPS  & unknown \\ \hline
    Fitbit & Dedicated hardware and Mobile App & Dedicated Hardware & 1M - 5M\cite{fitbit_store}\\ \hline
  \end{tabular}
  \caption{Comparison of exercise gamification implementations from
    the Google Play store}
  \label{table:competitor_comparison}
\end{table}

\subsection{Charity Miles}
\label{sec:charity_miles}
Charity Miles encourages you to exercise by facilitating you to raise
awareness of charities and help
others as you exercise. For every mile that you run or walk \$0.25 is
donated to a charity of your choice, and for every mile you cycle \$0.10
is donated. The money currently comes from a pool of \$1,000,000
gathered by the parent company - so you as an end user don't have to pay
anything. 

The money you accrue during an exercise session is called your
``sponsorship'' and is only passed onto your charity when you share
your success to a social network (Facebook or
twitter)\cite{charitymiles_terms}. Charity Miles say this is to
increase awareness of the charity, but it is also a good advertising
tool for them.

The incentive here is clear - run to donate. In essence, you are both
being paid to exercise and exercising to help out a charity. The
monetary backing allows you to quantify your exercise in terms you are
very familiar with whilst generating a feeling of satisfaction by
helping a charity.

The main drawback here is that there is no defined behaviour for what
happens when the \$1,000,000 pool runs out. From a business
perspective, the app could easily be re-purposed for non-centralised
pool funding where the end user could find local sponsorship. However
this entirely changes the user experience and explicitly requires
effort on the users side to gather this sponsorship. 

\subsection{Zombies, Run!}
\label{sec:zombies}
Zombies, Run! is a companion app that encourages you to exercise through
audio cues, rich story and a base-building game. As you run you are
being ``chased'' by a hoard of zombies that get closer and further
away from you as you exercise - the closer they are to catching you
the louder they become. 

In the app you have a base from where you defend yourself from the
zombie hoard and are required to leave the base to gather
supplies. The story narrative tells you what you need to gather and
you find these supplies based on the distance you travel and the route
you take. Your success is measured by the fortifications of your base,
which you can improve by running faster/further than before.

You are required to wear headphones and carry your mobile phone while
exercising. The headphones give you audio queues to immerse you in the
story, but it is a requirement of the game experience that you listen
to the soundscape provided - you cannot listen to your own music. As
endearing as groaning zombies may be, it is understandable that users
may tire of hearing it. Also, a user may be used to listening to their
own music and use that to zone out of the exercise so some users might
be reluctant to adopt this.


\subsection{WalkJogRun}
\label{sec:walkjogrun}
WalkJogRun helps you discover new routes that are geographically
close to where you are and record your favourite routes to share with
your friends. From your current location it finds and
proposes new exercise routes given a certain route length. It can
provide turn-by-turn direction help when you are exercising to ensure
you do not get lost and will also record an exact route as you run so
you can create a new route with the minimum of effort. 

Your workout is saved and can be later analysed so you can accurately
monitor your progress.

Since the routes are user provided, there is no guarantee that the
routes are safely traversable or applicable to all types of exercise -
a route that can be walked may not be applicable for a runner due to
terrain type, for example. 

There is no clear gamified experience in this application, so the
incentive is self-provided. The user can only use this to improve
their exercise experience as there is not much in the way of
encouragement. 

\subsection{Fitbit}
\label{sec:fitbit}
I have included a comparison with Fitbit products to display other
mediums in which this application can be used. Although it is not
mainly based around a mobile phone use it is an interesting comparison.

Fitbit provide a collection of dedicated hardware devices capable of
accurately measuring your distance travelled (and in some models the
height you have ascended in terms of ``flights of stairs'') as a means
of tracking your activity. At the core, it is a glorified pedometer
with web and phone apps to accompany it and display progress. The
devices and apps also make a prediction about the calories you burn each
day based upon the distance you travel when wearing the device, your
weight and stride length. 

Each device can show you the progress toward your daily goal (either
steps taken or predicted calorie burn) with various visualisation
techniques - the simplest of which is a row of five LED lights that
fill up as you near your goal. Synchronisation with compatible phones,
or through a computer if your device is not compatible, allows a more
in depth visualisation of your progress.

The gamification aspect is inherent in the design of these products -
you set your own goal and you can easily view your progress at a
glance. Upon completion of your goal the device vibrates and lights up
to notify you. You are also awarded badges for daily completion such
as walking 10,000 steps a day and lifetime completion goals such as
travelling 100 miles in total.

There is a clear downside to this - you are required to purchase a
bespoke device to fully utilise the experience provided by
Fitbit. When compared to an app that may be \$1.99 or thereabouts it is
a much larger start up cost than the previous examples. However this
larger start up cost may be an encouragement unto itself: the user may
feel obligated to commit to using the device since they are more
financially invested, which may help with the on boarding required to
fully benefit from these systems.

Most of these devices are reasonably discreet, the more advanced the
devices get the larger they become. Since these devices must be worn
mainly on the wrist there are social implications involved. Does the
user want everyone to know they are using a fitness tracker? The other
applications are more discreet by design, since they are a mobile app
they can be hidden easier. 

\section{Consideration of competitors concepts in this project
  implementation}
\label{sec:consideration}
Each of the above products utilise gamification in an effective way
and each implementation differs in how they utilise game
elements. This project builds upon some of the design decisions these
competitors have used, aiming to remove some of issues that make the
competitors products either unsustainable or flawed.

First and foremost, this project should not require any form of
additional hardware for the end user or monetary investment. A key
demographic that we are targeting are those users who do not exercise
outside often, so we do not want to create a barrier between those
users and them achieving their exercise goal. This is in reference to
the Fitbit product discussed in Section \ref{sec:fitbit} which
requires a dedicated hardware wristband, and the ``Zombies, Run!'' product
(Section \ref{sec:zombies}) which requires the user to purchase the
mobile app. Both of these could be reasons for an under encouraged
user to not take the first step to starting the exercise process.

This project should also not require a social aspect to achieve the
core goals for this application, unlike WalkJogRun (Section
\ref{sec:walkjogrun}) which has a heavy reliance on social
contributions from users in your area. Although WalkJogRun does not
provide a gamified experience, it does enhance your exercise
experience by showing you new running routes. The routes that are
shown as suggestions are entered into the system by other users as
they run. An analysis of crowdsourcing gamified content for exercise
encouragement is outwith the scope of this project due to the number
of players and the length of time required to complete a worthwhile
analysis. It would be interesting future work in the context of this
application, however. 

Charity Miles (Section \ref{sec:charity_miles}) also has a social
restriction where the user is required to share to a social networking
site information about the exercise session they have completed before
their charity of choice can receive the money the user has raised. It
can be argued either way as to whether or not this will be an extra
encouragement to the user, but this extra step is an obstacle
between the user and their complete success metric. This project
should endeavour to remove these obstacles that are not related to
their own personal ability: that is to say the user should only be
restricted access to their success metric because of the users effort
did not merit access (a user was not awarded a gold medal because the
user did not deserve a gold medal, not because they did not share
their running time to Facebook).

There should be no limitation to the app because of a reliance on an
outside party. In \emph{Charity Miles} there is a central reserve of
money which you can allocate to your charity of choice by exercising
but there is no defined behaviour for when this money runs out. This
central reserve of money is raised through the parent company securing
sponsorship and so there is a reliance on an outside party to maintain
the game experience. Reliance on outside parties is an inevitability
for small starting businesses but we should not allow this reliance to
affect the game experience for the user.










