% This example An LaTeX document showing how to use the l3proj class to
% write your report. Use pdflatex and bibtex to process the file, creating 
% a PDF file as output (there is no need to use dvips when using pdflatex).

% Modified 

\documentclass{l4proj}
\usepackage{url}
\usepackage{color}
\usepackage{graphics,graphicx}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{ifthen}
\usepackage{float}
\usepackage[superscript,biblabel]{cite}
\usepackage{hyperref}
\usepackage{mathtools}

\begin{document}

\newcommand{\todo}[1]{\textcolor{red}{#1}}
	
	
\title{Walk Around The World}
\author{Ryan Wells}
\date{\today}
\maketitle
\begin{abstract}
For many, cardiovascular exercise is an activity that we do not get
enough of\cite{exercise} . Studies have shown that physical exercise
between 1 hour a day for children over five years of age and as
little as 2.5 hours a week for adults has a huge positive impact on
our physical and emotional wellbeing\cite{govsurvey, amsurvey} .

This project aims to encourage users to do oudoor
cardiovascular exercise by positive encouragement through a Mobile
Application. By tracking the distance the user travels in each
exercise session through device GPS and translating this into
well-known outdoor pursuits allows the user to quantify their exercise
in terms of well known physical and cultural achievements. These
achievements will take badge-like form and other metrics of success
such as leaderboards will also exist. 

Outdoor pursuits will contain: actual routes such as the West Highland
Way and the climb to the top of Everest; popular culture references
such as ``Route 66'' and the (approximate) distance that \emph{Frodo}
and \emph{Sam} travelled in J.R.R.Tolkein's ``The Lord of the Rings'';
and Global distances such as the distance between capitol cities and
simple metrics such as the first one hundred miles. 

The user will exercise with a compatible device on their
person. During exercise the device will track and log the distance
that the user has travelled and add this to their accumulators. If
this device is their smart phone then the Mobile Application will
notify them immediately when they have completed an outdoor pursuit,
otherwise they will be notified when data is collected from the
dedicated hardware.

This application will encourage social exercise through outdoor
pursuits that can only be completed through teamwork with other users
that have met in real life. This intends to extend the influence of
rewarding exercise by encouraging users to include their peer group in
exercising with them and sharing their successes. For a user,
this is incentivised by unlocking achievements only achievable through
this social interaction.

An analysis of user trends such as: frequency and duration of exercise
with regards to duration of application use; change in exercise
duration when achievements are awarded during exercise; and quantity of
social interaction with regards to application use life cycle will be
used to indicate whether or not gamification techniques could have a
valid application in cardiovascular exercise. 

\end{abstract}
%\educationalconsent
\tableofcontents
%==============================================================================
\pagebreak
\pagenumbering{arabic}
\input{introduction}
\input{related_work}
\input{method}
\input{results}
\input{conclusion}

\section{Acknowledgements}
I would like to thank ...




\bibliographystyle{unsrt}
\bibliography{wells-2013-urbanexplorer}
\end{document}
